### 2012-12-05 inst/examples/mcmc-gp.Rmd
Notes on `tgp` package MCMC

```{r results="hide"}
knit("beverton_holt_data.Rmd")
true <- data.frame(x = X, y = sapply(X, f, 0, p))
```

```{r}
gp <- bgp(X=obs$x, XX=X, Z=obs$y)
```

@@

```{r}
center <- tgp:::tgp.choose.center(gp, "mean")
Z.mean <- center$Z
X_vals <- center$X[, 1]
o <- order(X_vals)
Zb.q1 <- Z.mean + 1.96 * sqrt(c(gp$Zp.ks2, gp$ZZ.ks2))
Zb.q2 <- Z.mean - 1.96 * sqrt(c(gp$Zp.ks2, gp$ZZ.ks2))
V <- c(gp$Zp.ks2, gp$ZZ.ks2)[o]
dat <- data.frame(x=X_vals[o], y=Z.mean[o], ymin=Zb.q1[o], ymax=Zb.q2[o])
```

>>>>

```{r results="hide"}
knit("~/Documents/code/nonparametric-bayes/inst/examples/beverton_holt_data.Rmd")
true <- data.frame(x = x_grid, y = sapply(x_grid, f, 0, p))
XX <- as.matrix(x_grid)
X <- as.matrix(obs$x)
Z <- as.matrix(obs$y)
```

## Call to `tgp` function `bgp()`

Tweak the hyperparameter prior's parameters (`s2.p`) to influence the GP inferred.

```{r}
gp <- bgp(X=X, XX=XX, Z=Z, meanfn="constant", bprior="b0", BTE=c(1000,6000,2), m0r1=TRUE, verb=4, corr="exp", trace=TRUE, s2.p=c(5,10), tau2.p=c(5,10), s2.lam="fixed", tau2.lam="fixed")
```

## parsing the function call

* `X` is the observed X values (or matrix of appropriate dimension)
* `XX` the desired predicted grid. (will also predict on X)
* `Z` is observed Y values (or matrix of appropriate dimension)
* `m0r1` means scale data to mean zero and unit range
* `bprior` is the prior distribution, defaulting to `bfalt`. `b0` is a Gaussian.  This prior is for the Gaussian process and is not to be confused with the hyperpriors describing the distribution of various parameters.
* `BTE` Burn-in B steps, Terminate after T steps, sample Every E steps. 
* `verb` Very verbose output
* Other options (for adaptive/query learning, etc, igored for now.  Focus on priors).  
* The priors for the hyperparameters are given by `s2.p`, `tau.p` etc (both are Inverse Gammas).  Their parameters can vary hiearchically in general, but we hold them fixed, `s2.lam = 'fixed'`, etc.)

## Hyperparameters

Hyperparameter distributions returned in the trace:

```{r}
names(gp$trace$XX[[1]][2500,])
```

* `s2` is $\sigma^2$ in Gramacy and $\tau$ in Rassmussen and Williams, is the scale of the noise. Comes from an inverse gamma prior.  Set by inverse-gamma parameters `a0` and `g0`.

* `tau2` Also comes from an inverse gamma prior. It is also set by inverse-gamma parameters `a0` and `g0`.

$$Z | \beta, \sigma^2, K \sim N_n(\mathbf{F} \beta, \sigma^2 \mathbf{K}) $$
$$ \beta | \sigma^2, \tau^2, \mathbf{W}, \beta_0 ~ N_m(\beta_0, \sigma^2 \tau^2 \mathbf{W} )  $$
$$\beta_0 | N_m(\mathbf{\mu}, \mathbf{B})$$

The parameters in this run are constant across the trace points (because hyperparameters are fixed).

```{r}
gp$trace$XX[[1]][2500,]
gp$trace$XX[[3]][2500,]
```

```{r}
pars <- melt(gp$trace$XX[[1]], id = "index")
ggplot(pars, aes(index, value)) + geom_line() + facet_wrap(~variable, scales="free_y")
````

It is possible to calculate summary statistics and check out the distribution of these fellows.

```{r}
mean(gp$trace$XX[[1]][,"nug"])
ggplot(pars, aes(value)) + geom_histogram() + facet_wrap(~variable, scales="free")
```

* We keep the heirachical priors fixed in the `tau.lm`, `s2.lm`, etc. `nug` fixed by default.  Note this results in the the priors from which these are sampled are fixed: e.g. `gp$trace$hier$s2.a0`, `gp$trace$hier$s2.g0`

* `nug.p` This is the "nugget," the measurement error.  Comes from a mixture of gamma prior parameter (initial values) for the nugget parameter `c(a1,g1,a2,g2)` where `g1` and `g2` are scale (`1/rate`) parameters. _Default reduces to simple exponential prior._ Specifying `nug.p = 0` fixes the nugget parameter to the "starting" value in `gd[1]`, i.e., it is excluded from the MCMC

* `index` is the sampling point (from `BTE`, we see we sample starting at step 1000 and ending at step 6000, recording every 2 steps, so there are $(T-B)/E = 2500$ index points).  
* `lambda` possibly a mixing parameter in the importance sampler temperature??
* `beta0` No mention in the manual
* `d` The parameter for priors using the hierachical exponential distribution for the parameters `a1`, `a2`, `g1`, `g2`.  
* `b` ??

* Where are the hyperparameters for the correlation function (e.g. the length-scale for the Gaussian??)

## Extracting the estimated process

(The **MAP**, maximum a posteriori expected mean and variance).

@@

```{r}
V <- gp$ZZ.ks2
dat <- data.frame(x   = gp$XX[[1]], 
                 y    = gp$ZZ.km, 
                 ymin = gp$ZZ.km - 1.96 * sqrt(gp$ZZ.ks2), 
                 ymax = gp$ZZ.km + 1.96 * sqrt(gp$ZZ.ks2))
```
